CYO Project Report - Brazilian Ecommerce

### 2021-07-11 finalReport.Rmd
added target encoding for seller_id and feature scaling for knn

```{r Install or load packages,include=FALSE}
# TODO use conditional loading or installation of library
# Remove plotly related code for pdfs
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
if (!require('caret')) install.packages('caret'); library('caret')
if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
if (!require('plotly')) install.packages('plotly'); library('plotly')
if (!require('gam')) install.packages('gam'); library('gam')
```

@@

For products data set,
- we read data, remove unnecessary fields, join with english names whereever available
- remove products where dimension and related attributes are not available
- assign default category name of UNCATEGORIZED for products where same was not available
```{r Data wrangling-products data set, include=FALSE}
products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)

@@

# assign category name for blank category
products_dataset <- products_dataset %>% mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
```
The wrangled data looks like below:
```{r Data wrangling result-products data set, echo=FALSE}
products_dataset %>% head()
```

For geolocation data set
```{r Data wrangling- geolocation data set, echo=FALSE}
# geo location details
# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
# it can have geolocation_zip_code_prefix shared across multiple places
# refer: https://codigo-postal.org/en-us/brazil/
# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil. The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:

@@

```{r Data wrangling-customers data set, echo=FALSE}
# customers data set  with zip code prefix and can be used to uniquely identify a customer
# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
# customer_zip_code_prefix --> first five digits of customer zip code
customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))

@@

```{r Data wrangling-sellers data set, echo=FALSE}
# seller data set with zip code prefix
# seller_zip_code_prefix --> first 5 digits of seller zip code
sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
# TODO review if we need it
# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))

@@

Wrangled customers dataset
- removed data where delivery was not made
- removed data which was delivered but all the delivery details were not available
```{r Data wrangling-orders data set, echo=FALSE}
# order data set with various delivery related times
# note each customer_id is unique per order
# order_purchase_timestamp -> Shows the purchase timestamp.
# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% rename(payment_approved_at=order_approved_at,handed_over_to_carrier_date=order_delivered_carrier_date)  %>% mutate(order_status=as.factor(order_status),order_purchase_timestamp=as_datetime(order_purchase_timestamp),payment_approved_at=as_datetime(payment_approved_at),handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),order_delivered_customer_date=as_datetime(order_delivered_customer_date),order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))

@@

```{r Data wrangling-order items data set, echo=FALSE}
# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))

@@

Filter out data where sellers location details are not available
```{r Refine sellers dataset, include=FALSE}
#
# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# These all need to be handled in data wrangling

@@

Wrangled sellers dataset ,looks like below:
```{r Refined sellers dataset result, echo=FALSE}
sellers_dataset %>% head()
```

```{r Refine customers data set, include=FALSE}
# similarly let see customers for whom we don't have data in geolocation_dataset
customers_to_ignore <- customers_dataset %>% anti_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix'))

@@

```{r Refined customers dataset result, echo=FALSE}
customers_dataset %>% head()
```

@@

```{r Combine data sets, include=FALSE}
# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()

# let's combine with seller details
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()

# let's combine with unique customer and its location dataset
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()

# let's combine with products data as well
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name)) %>%
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
```

Calculating current RMSE range
```{r Current RMSE, echo = FALSE}
# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10

@@

# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
# it is possible some products are bulky or require differen mode of transport like ship 
p1 <- data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')

library(plotly)
ggplotly(p1)

rm(data_for_rmse)
```

```{r Prepare final wrangled data set, echo=FALSE}
# we will predict number of hours to deliver based on various predictors. This seems more appropriate. Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
wrangled_data <- raw_data %>%
  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
  select(-order_purchase_timestamp,-order_delivered_customer_date,-order_estimated_delivery_date)
```

@@

wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars) %>% select(-product_category_name)

@@

```{r Final wrangled data,echo=FALSE}
wrangled_data %>% head()
```

```{r Cleanup and save required dataset}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
```

@@

```{r Method to derive delay in handover to carrier, include = FALSE}
# calculated in hours
get_delay_in_handover_to_carrier_per_seller_id <- function(df) {
  df %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600)  %>% 
    group_by(seller_id) %>% 
    summarise(mean_penalty_to_consider=mean(delay))
}
```

@@

Split data into training and validation set, where validation set will be used later
```{r Split data into training and validation set, include = FALSE}
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)

@@

# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name') %>% semi_join(training_dataset,by='seller_id')
validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
# test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')

@@

```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
# TODO review it, fix it as required
training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```

@@

```{r Prepare training and test data for ,include=FALSE}
# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)

test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
```

```{r knn model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')

@@

```{r knn model data scaline under test, echo=FALSE}
# TODO review and improve or remove this as required
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

tuningControl=trainControl(method = "cv", number = 3, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#194.8233 for k=34

@@

```{r GamLoess model with scaling, echo = FALSE}
# TODO review-keep or remove it
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

@@

Data used maybe incorrect
Review joins and if multiple orders of same type are being delivered at the same time
refer
> training_dataset %>% filter(number_of_hours>=354.73 & number_of_hours<354.74)
   price freight_value product_category_name product_weight_g number_of_hours volume
1   59.0         13.43          garden_tools             1550        354.7386  19800
2   59.0         13.43          garden_tools             1550        354.7386  19800
3   59.0         13.43          garden_tools             1550        354.7386  19800
4   59.0         13.43          garden_tools             1550        354.7386  19800
5   59.0         13.43          garden_tools             1550        354.7386  19800
6   59.0         13.43          garden_tools             1550        354.7386  19800
7   59.0         13.43          garden_tools             1550        354.7386  19800
8   59.0         13.43          garden_tools             1550        354.7386  19800
9   59.0         13.43          garden_tools             1550        354.7386  19800
10  59.0         13.43          garden_tools             1550        354.7386  19800
11  59.0         13.43          garden_tools             1550        354.7386  19800
12  66.0         14.21       furniture_decor             1150        354.7319  28175
13  46.9         17.63        sports_leisure              300        354.7378   2304
   distance_between_two_points mean_penalty_to_consider
1                       469.74                -51.60514
2                       469.74                -51.60514
3                       469.74                -51.60514
4                       469.74                -51.60514
5                       469.74                -51.60514
6                       469.74                -51.60514
7                       469.74                -51.60514
8                       469.74                -51.60514
9                       469.74                -51.60514
10                      469.74                -51.60514
11                      469.74                -51.60514
12                      807.32                -56.57523
13                     1083.46               -147.48749

>>>>

```{r Install or load packages,include=FALSE, echo = FALSE}
# TODO use conditional loading or installation of library
# Remove plotly related code for pdfs
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
if (!require('caret')) install.packages('caret'); library('caret')
if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
if (!require('knitr')) install.packages('knitr'); library('knitr')
```

@@

For products data set,
- we read data, remove unnecessary fields, join with english names whereever available
- remove products where dimension and related attributes are not available
- assign default category name of UNCATEGORIZED for products where same was not available
```{r Data wrangling-products data set, include=FALSE, echo=FALSE}
products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% select(-product_name_lenght,-product_description_lenght,-product_photos_qty)

@@

# assign category name for blank category
products_dataset <- products_dataset %>% mutate(product_category_name=if_else(product_category_name=='','UNCATEGORIZED',product_category_name))
```
The wrangled data looks like below:
```{r Data wrangling result-products data set, echo=FALSE}
products_dataset %>% head() %>% knitr::kable()
```

For geolocation data set
```{r Data wrangling- geolocation data set, include=FALSE, echo=FALSE}
# geo location details
# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
# it can have geolocation_zip_code_prefix shared across multiple places
# refer: https://codigo-postal.org/en-us/brazil/
# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil. The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:

@@

```{r Data wrangling-customers data set, include=FALSE, echo=FALSE}
# customers data set  with zip code prefix and can be used to uniquely identify a customer
# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
# customer_zip_code_prefix --> first five digits of customer zip code
customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
# %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))

@@

```{r Data wrangling-sellers data set, echo=FALSE, include=FALSE}
# seller data set with zip code prefix
# seller_zip_code_prefix --> first 5 digits of seller zip code
sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
# TODO review if we need it
# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))

@@

Wrangled customers dataset
- removed data where delivery was not made
- removed data which was delivered but all the delivery details were not available
```{r Data wrangling-orders data set, include=FALSE, echo=FALSE}
# order data set with various delivery related times
# note each customer_id is unique per order
# order_purchase_timestamp -> Shows the purchase timestamp.
# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% 
  rename(payment_approved_at=order_approved_at,
         handed_over_to_carrier_date=order_delivered_carrier_date)  %>% 
  mutate(order_status=as.factor(order_status),
         order_purchase_timestamp=as_datetime(order_purchase_timestamp),
         payment_approved_at=as_datetime(payment_approved_at),
         handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),
         order_delivered_customer_date=as_datetime(order_delivered_customer_date),
         order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))

@@

```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))

@@

Filter out data where sellers location details are not available
```{r Refine sellers dataset, include=FALSE, echo=FALSE}
#
# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# These all need to be handled in data wrangling

@@

Wrangled sellers dataset ,looks like below:
```{r Refined sellers dataset result, echo=FALSE}
sellers_dataset %>% head() %>% knitr::kable()
```

```{r Refine customers data set, include=FALSE, echo=FALSE}
# similarly let see customers for whom we don't have data in geolocation_dataset
customers_to_ignore <- customers_dataset %>% anti_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix'))

@@

```{r Refined customers dataset result, echo=FALSE}
customers_dataset %>% head() %>% knitr::kable()
```

@@

```{r Combine data sets, include=FALSE, echo=FALSE}
# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()

# let's combine with seller details
# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()

# let's combine with unique customer and its location dataset
# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()

# let's combine with products data as well
# orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name)) %>%
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
#  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
```

Calculating current RMSE range
```{r Current RMSE, include=FALSE, echo = FALSE}
# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10

@@

# if we look at graph we can  see there is huge difference between estimated data and actual delivery date
# it is possible some products are bulky or require differen mode of transport like ship 
data_for_rmse %>% mutate(diff=estimated_time_from_order_to_delivery-1/2*(time_taken_from_order_to_delivery_rangemax+time_taken_from_order_to_delivery_rangemin)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 24*2) + xlab('Estimation error where each bar represents 2 days')

rm(data_for_rmse)
```
```{r Range of current rmse, echo=FALSE}
data.frame(type=character(),value=numeric()) %>% 
  add_row(type='RMSE Max',value=RMSE_value_based_on_max) %>% 
  add_row(type='RMSE Min',value=RMSE_value_based_on_min) %>% 
  knitr::kable(caption='Current RMSE')
```

Final wrangled data looks like below:
```{r Prepare final wrangled data set, include=FALSE}
# we will predict number of hours to deliver based on various predictors. This seems more appropriate. Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
wrangled_data <- raw_data %>%
  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
  select(-order_purchase_timestamp,-order_delivered_customer_date,-order_estimated_delivery_date)
```

@@

wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars)

@@

```{r Final wrangled data,echo=FALSE}
wrangled_data %>% head() %>% knitr::kable()
```

```{r Cleanup and save required dataset, include=FALSE, echo=FALSE}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
```

TODO
Use target encoding for categorical variables
Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.

@@

```{r Method to derive delay in handover to carrier, include = FALSE}
# calculated in hours
get_delay_in_handover_to_carrier_per_seller_id <- function(df) {
  df %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600)  %>% 
    group_by(seller_id) %>% 
    summarise(mean_delay_in_handover=mean(delay))
}
```

@@

Split data into training and validation set, where validation set will be used later
```{r Split data into training and validation set, include = FALSE, echo = FALSE}
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)

@@

# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
validation_dataset <- validation_dataset %>%
  semi_join(training_dataset,by='product_category_name') %>%
  semi_join(training_dataset,by='seller_id')
# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
#test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')

@@

```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```

@@

```{r Prepare training and test data for ,include=FALSE}
# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)

test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date)
```

```{r knn model, echo = FALSE}
#TODO can be removed
set.seed(seed=1,sample.kind = 'Rounding')

@@

```{r Apply Target encoding, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:30000,]
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_subset_for_encoding$number_of_hours,p=0.2,times=3,list=FALSE)

  weights <- seq(0,6,3)

  weight_to_use_per_partition <- sapply(1:3, function(index) {
  #lambda_data_set <- prepare_lambda_data_set(index)
  #train_lambda_set <- lambda_data_set[[1]]
  #test_lambda_set <- lambda_data_set[[2]]  
  train_weight_set <- train_subset_for_encoding[-test_weight_indices[,index],]
  test_weight_set <- train_subset_for_encoding[test_weight_indices[,index],]

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_id')

  weight_rmses <- sapply(weights, function(weight) {
      seller_id_effects <- train_weight_set %>% 
        group_by(seller_id) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

      train_weight_set <- train_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)

      test_weight_set <- test_weight_set %>% left_join(seller_id_effects,by='seller_id') %>% rename(seller_id_effect=mean) %>% select(-product_category_name,-seller_id)

      set.seed(seed=1,sample.kind = 'Rounding')
      preProcessScalingValues <- preProcess(train_weight_set %>% select(-number_of_hours), method = c("center", "scale"))

      scaled_train_subset <- predict(preProcessScalingValues, train_weight_set)
      scaled_test_subset <- predict(preProcessScalingValues, test_weight_set)

      # TODO remove below scaling was erroneous. train scaling needs to be used for test data
      # below is just demo of how to use scale
      # scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
      # scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

      # TODO can we skip modelling while evaluating weight to use or for target encoding?

      tuningControl=trainControl(method = "cv", number = 2, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

      knn_fit_for_weight

      predictions <- predict(knn_fit_for_weight,scaled_test_subset %>% select(-number_of_hours))

      RMSE(scaled_test_subset$number_of_hours,predictions) 
  })

  # selecting weight leading to lowest RMSE
  weight_to_use <- weights[which.min(weight_rmses)]
  paste('Weight to use: ', weight_to_use, ', corresponding RMSE: ', weight_rmses[which.min(weight_rmses)])

  print(weight_rmses)
  weight_to_use
})

weight_to_use_per_partition

weight_to_use <- mean(weight_to_use_per_partition)  
```

```{r Post regularization, echo=FALSE}

# prepare seller_if_effect based on regularisation weight we have identified
  global_mean <- mean(train_subset$number_of_hours)

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)

seller_id_effects_list <- lapply(1:5, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

    seller_id_effects <- train_weight_set %>% 
      group_by(seller_id) %>%
      summarise(mean=mean(number_of_hours),n=n()) %>%
      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
      select(-n) %>% rename()

    seller_id_effects
  }
)

sie <- seller_id_effects_list[[1]]
for (index in 2:5) {
   sie <- sie %>% full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
}

# whereever value is NA, we have used global mean
sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))

# as different partitions may not have exactly same data set combined the results
seller_id_effects <- tibble(sie[,1], seller_id_effect=rowMeans(sie[,-1],na.rm = TRUE))

train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))

test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id)
```

```{r knn model data scaled under test, echo=FALSE}
# TODO review and improve or remove this as required
set.seed(seed=1,sample.kind = 'Rounding')
preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))

scaled_train_subset <- predict(preProcessScalingValues, train_subset)
scaled_test_subset <- predict(preProcessScalingValues, test_subset)

# TODO remove below scaling was erroneous. train scaling needs to be used for test data
# below is just demo of how to use scale
# scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
# scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

tuningControl=trainControl(method = "cv", number = 3, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#192.8755 for k=27

@@

```{r GamLoess model with scaling, echo = FALSE}
# TODO review-keep or remove it
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

@@

Data used maybe incorrect
Review joins and if multiple orders of same type are being delivered at the same time
refer
> training_dataset %>% filter(number_of_hours>=354.73 & number_of_hours<354.74)
   price freight_value product_category_name product_weight_g number_of_hours volume
1   59.0         13.43          garden_tools             1550        354.7386  19800
2   59.0         13.43          garden_tools             1550        354.7386  19800
3   59.0         13.43          garden_tools             1550        354.7386  19800
4   59.0         13.43          garden_tools             1550        354.7386  19800
5   59.0         13.43          garden_tools             1550        354.7386  19800
6   59.0         13.43          garden_tools             1550        354.7386  19800
7   59.0         13.43          garden_tools             1550        354.7386  19800
8   59.0         13.43          garden_tools             1550        354.7386  19800
9   59.0         13.43          garden_tools             1550        354.7386  19800
10  59.0         13.43          garden_tools             1550        354.7386  19800
11  59.0         13.43          garden_tools             1550        354.7386  19800
12  66.0         14.21       furniture_decor             1150        354.7319  28175
13  46.9         17.63        sports_leisure              300        354.7378   2304
   distance_between_two_points mean_delay_in_handover
1                       469.74                -51.60514
2                       469.74                -51.60514
3                       469.74                -51.60514
4                       469.74                -51.60514
5                       469.74                -51.60514
6                       469.74                -51.60514
7                       469.74                -51.60514
8                       469.74                -51.60514
9                       469.74                -51.60514
10                      469.74                -51.60514
11                      469.74                -51.60514
12                      807.32                -56.57523
13                     1083.46               -147.48749
